\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage[english]{babel}
\usepackage{listings}
\usepackage{booktabs}
\usepackage{hyperref}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  keywordstyle=\bfseries,
  commentstyle=\color{gray}\itshape,
  showstringspaces=false,
  tabsize=4,
  morekeywords={func,type,struct,chan,go,select,case,return,if,for,range,
  make,nil,true,false,var,const,package,import,defer,error},
}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Distributed Configuration Updater:\\
  Design and Implementation of a CRDT-Based\\
Settings Replication System}

\author{\IEEEauthorblockN{Gabriele Da Re}
  \IEEEauthorblockA{\textit{Dept. Mathematics} \\
    \textit{Universit\`{a} degli Studi di Padova}\\
    Padova, Italy \\
  gabriele.dare00@gmail.com}
}

\maketitle

\begin{abstract}
  This paper describes the architecture, implementation and decisions made during
  the creation of a configuration synchronization system for small server clusters.
  The system aims to replicate Key-Value (KV) entries across nodes through: Last-Writer-Wins (LWW),
  Conflict-free Replicated Data Type (CRDT), gRPC transport protocol and a Hybrid Logical Clock (HLC) for causal ordering.
  We explain the decisions made designing each module and the problems encountered/mitigated during development such as: Clock-skew,
  crash persistance handling, tombstones management and the anti-entropy approach and lifecycle enforcement between initialisation
  and runtime phases.
\end{abstract}

\begin{IEEEkeywords}
  CRDT, distributed systems, data consistency, hybrid logical clocks,
  gRPC, Go, key-value store
\end{IEEEkeywords}

% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}
% ─────────────────────────────────────────────────────────────────────────────

Configuration management across a server cluster is a common operational
challenge. Changes to settings such as feature flags, service endpoints,
parameteers must propagate to every node without a coordinator becoming
the single failure point.

Approaches like \href{https://etcd.io/}{etcd},
\href{https://developer.hashicorp.com/consul}{Consul} rely on a leader node or a
distributed database. These provide linearisability at the cost of availability
(CP part of the CAP theorem and ACID principle): when the leader or a majority quorum is unreachable,
there is no working system. For configuration data, human operators can intervene
on conflicts and having stale data in some circumstances is relatively low risk,
thus the trade-off favours Availability and Partition tolerance having \textbf{eventual} Consistency (AP).

Conflict-free Replicated Data Types (CRDTs) \cite{shapiro2011} offer an
alternative: every node can accept writes locally, the data
structure guarantees that any two nodes with the same set of updates will
converge to the same state, without coordination. This paper describes a
practical implementation of a CRDT based configuration store targeting small,
fixed-membership clusters where $O(100)$ configuration keys are replicated
across $O(10)$ nodes. This limit can be pushed up to $O(1000)$ keys and $O(50)$ nodes
if necessary, based upon limits imposed from the full snapshot, sync implementation \ref{sec:antientropy}.

The implementation is written in Go and uses gRPC for inter-node
communication. The remainder of this paper is structured as follows.
Section~\ref{sec:arch} presents the overall system architecture.
Section~\ref{sec:crdt} describes the CRDT data model.
Section~\ref{sec:hlc} covers the Hybrid Logical Clock.
Section~\ref{sec:concurrency} explains the concurrency model.
Section~\ref{sec:network} details the network layer.
Section~\ref{sec:antientropy} discusses anti-entropy reconciliation.
Section~\ref{sec:persistence} describes crash-safe persistence.
Section~\ref{sec:gc} presents tombstone garbage collection.
Section~\ref{sec:lifecycle} covers lifecycle and error handling.
Section~\ref{sec:testing} outlines the testing strategy.
Section~\ref{sec:conclusion} concludes.

% ─────────────────────────────────────────────────────────────────────────────
\section{System Architecture}
\label{sec:arch}
% ─────────────────────────────────────────────────────────────────────────────

The system is structured as five Go packages with a strict dependency
hierarchy, shown in Table~\ref{tab:packages}.

\begin{table}[htbp]
  \caption{Package responsibilities}
  \label{tab:packages}
  \begin{center}
    \begin{tabular}{ll}
      \toprule
      \textbf{Package} & \textbf{Responsibility} \\
      \midrule
      \texttt{types}      & Shared domain types: HLC, SettingEntry, Snapshot \\
      \texttt{crdt}       & LWW CRDT, state machine, persistence \\
      \texttt{logic}      & Local settings file watcher and writer \\
      \texttt{network}    & gRPC server, streaming client, proto conversion \\
      \texttt{controller} & Wires all packages, lifecycle and backgrounds tasks \\
      \midrule
      \texttt{main}       & CLI entry-point \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

The data flows through the system following four paths, wired up in
\texttt{controller.Run}:

\begin{enumerate}
  \item \textbf{Local write}: the \texttt{logic} watcher detects file
    changes on write and pushes a \texttt{SettingEntry} to \texttt{crdt.NotifyLocal},
    which queues it in local channel of the CRDT actor. The CRDT ticks
    its clock, timestamps the entry and merges it, persists state, and calls
    \texttt{OnBroadcast} delegating the broadcast to the network layer.
  \item \textbf{Remote write}: a peer to which the node is subscribed streams an update,
    \texttt{network.Client} calls \texttt{crdt.NotifyRemote}. The CRDT
    updates its clock, merges, persists, and calls \texttt{OnFileSync}
    that delegates the \texttt{logic} layer to write the change locally.
  \item \textbf{Anti-entropy}: a ticker in the controller calls \texttt{syncAllPeers},
    which sends the local snapshot to each peer via a Sync RPC whose returns (if any)
    any newer entries through \texttt{NotifyRemote}.
  \item \textbf{Tombstone GC}: another separate ticker calls
    \texttt{crdt.PurgeTombstones}, whose adhibited at removing stale deleted entries
    from the in-memory state map via the CRDT actor model.
\end{enumerate}

The controller is the only package that depends on all the others, preventing import cycles.

% ─────────────────────────────────────────────────────────────────────────────
\section{CRDT Data Model}
\label{sec:crdt}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Data Model}

The data structure applies \textbf{LWW-Register} semantics \cite{shapiro2011}
per key in a dictionary: each \texttt{SettingEntry} carries a payload,
an HLC as timestamp and a tombstone flag. The tombstone turns each register into
a two-state register (live/deleted), deletion itself is a timestamped
write, subject to the same LWW rule.

\begin{lstlisting}
type SettingEntry struct {
    Key     string
    Value   string
    Clock   HLC
    Deleted bool
}
\end{lstlisting}

\subsection{Merge Semantics}
\label{merge-semantics}

Two nodes merge by iterating one map against the other and checking
for each key the entry with the higher clock thus the winning one. On equal clocks,
the lexicographically larger NodeID is used as tiebreaker,
ensuring deterministic total order.

\begin{lstlisting}
func (c *CRDT) merge(incoming SettingEntry) bool {
    existing, ok := c.state[incoming.Key]
    if ok && !existing.Clock.Before(incoming.Clock) {
        return false
    }
    c.state[incoming.Key] = incoming
    return true
}
\end{lstlisting}

\subsection{Tombstone Semantics}

Deletions are represented as tombstones: entry with \texttt{Deleted~=~true}
and a fresh clock. Tombstones undergo the same LWW merge as the payload and
can be displaced by a later write. This prevents the anomaly where a deletion
arrives after a concurrent write and erroneously removes the newer value.

\subsection{CvRDT}

The implementation is a hybrid CRDT\@: each local state change broadcasts a
\texttt{SettingEntry} to peers (operation-based style), while a timed
\texttt{Sync} provides anti-entropy reconciliation (state-based style).
Because LWW merge is commutative and idempotent, each broadcast is algebraically
equivalent to a partial state merge. Convergence is guaranteed
entirely by the state-based anti-entropy, the broadcast is a best-effort optimisation.

\subsection{Convergence Proof (CvRDT)}
\label{sec:cvrdtproof}

\paragraph{State forms a join-semilattice.}
Define the replica state as $S = \mathrm{Map}[\mathit{String} \to
\mathit{SettingEntry}]$.

\noindent Impose a partial order $\forall k \in Keys$
\begin{equation}
  S_1 \leq S_2 \iff \forall k: S_1[k].\mathit{Clock} \leq S_2[k].\mathit{Clock}
\end{equation}
where \texttt{Before} defines the strict order $<$ on clocks (Definition~\ref{sec:hlc}),
and $\leq$ is its reflexive closure: $a \leq b \iff a < b \;\lor\; a = b$.
The least upper bound per key is
\begin{equation}
  (S_1 \sqcup S_2)[k] = \arg\max_{\mathit{Clock}}\bigl(S_1[k],\; S_2[k]\bigr)
\end{equation}
which is exactly \texttt{merge()}.

\paragraph{Merge implements the LUB.}
Let $A$, $B$, $C$ be \texttt{SettingEntry} values for the same key.

\begin{itemize}
  \item \textbf{Commutative:} $\mathrm{merge}(A,B) = \mathrm{merge}(B,A)$, since
    $\max(\mathit{Clock}_A, \mathit{Clock}_B)$ is symmetric.
  \item \textbf{Idempotent:} $\mathrm{merge}(A,A) = A$, since
    $\max(T,T) = T$.
  \item \textbf{Associative:} $\mathrm{merge}(A,\mathrm{merge}(B,C)) = \mathrm{merge}(\mathrm{merge}(A,B),C)$, since $\max$ is associative.
\end{itemize}

\paragraph{Monotonicity.}
Every merge moves state upward in the lattice:
\begin{equation}
  S \leq \mathrm{merge}(S,\, e),
\end{equation}

since \texttt{merge} replaces an entry only when the incoming clock is strictly
greater, and never decreases a clock value.

\paragraph{Liveness.}
\texttt{syncAllPeers} runs periodically exchanging full snapshots. After a
network partition, reconnection triggers \texttt{Sync}, delivering all missing updates.
Hence every update eventually reaches every replica.

\paragraph{Conclusion.}
By the Eventual Consistency theorem for CvRDTs \cite{shapiro2011}: any
two replicas that have received the same set of updates hold identical maps,
regardless of delivery order.

% ─────────────────────────────────────────────────────────────────────────────
\section{Hybrid Logical Clock}
\label{sec:hlc}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Motivation}

A pure Lamport clock provides causal ordering but loses the physical-time
locality property: events that happen close together in real time may receive
widely separated clock values if logical increments accumulate. Hybrid Logical
Clocks \cite{kulkarni2014} integrates the Lamport clock with a wallTime,
preserving the benefits of both.

\subsection{Clock Structure}

Each HLC is a triple $(wall, logical, nodeID)$. The wall field stores
Unix nanoseconds, the logical field is a 32-bit logical counter, the NodeID is a UUID
string adhibited for the tiebreak. \ref{merge-semantics}. This ensures \texttt{Before()}
commutativity since $\forall n_i, n_j\in Nodes,\; i\neq j: Id(n_i) \neq Id(n_j)$

\subsection{Tick and Update Rules}

On a local event (\emph{tick}), the clock advances ($\ell_l, \ell_r$ for
$logical_{local}, logical_{received}$, $w_l, w_r$ for $wall_{local}, wall_{received}$):
\begin{align}
  wall' &= \max(w_l,\; now), \notag \\
  logical' &=
  \begin{cases}
    logical + 1 & wall' = w_l \\
    0           & otherwise
  \end{cases}
\end{align}

On receiving a remote clock (\emph{update}), the canonical
Kulkarni--Demirbas rule \cite{kulkarni2014} is applied:
\begin{equation}
  wall' = \max(w_l,\; w_r,\; now)
\end{equation}
The logical counter then captures concurrency at the chosen wall time
:
\begin{equation}
  logical' =
  \begin{cases}
    \max(\ell_l, \ell_r) + 1 & w' = w_l = w_r \\
    \ell_r + 1                 & w' = w_r \neq w_l \\
    0                          & otherwise
  \end{cases}
\end{equation}

\subsection{Clock Skew Handling}

\begin{lstlisting}
if received.WallTime - now > int64(10*time.Second) {
    return fmt.Errorf("rejecting: clock too far in future")
}
\end{lstlisting}

Rejection maps cleanly onto two failure cases.

\paragraph{Sender clock too far ahead} (e.g.\ B at $+90s$, threshold
$60s$): every correctly synchronised node rejects the entry.
B's writes are dropped cluster-wide until its NTP is corrected. No divergence occurs

\paragraph{Receiver clock too far behind} (e.g.\ A at $-30s$, B at
$+50s$): from A's perspective the skew is $80s > 60s$, so A rejects,
from a synchronised node C the skew is $50s < 60s$, so C accepts.
A and C temporarily diverge. This divergence is \emph{A's fault}: a
$30s$ NTP drift is an anomaly that requires intervention regardless
of this system. Once A's clock is corrected, the next anti-entropy
round delivers the missing entries and convergence is restored.
The divergence window is bounded by the NTP correction time, not permanent.

The $5s$ threshold is a configurable operational boundary. Nodes
whose NTP drift pass it, diverge until clock correction, this is
resolved at the infrastructure level (NTP monitoring,
alerting) rather than silently accepting future-stamped entries that would risk
clock poisoning across the cluster.

% ─────────────────────────────────────────────────────────────────────────────
\section{Concurrency Model}
\label{sec:concurrency}
% ─────────────────────────────────────────────────────────────────────────────

All mutations to the CRDT state map are serialised through a single actor
goroutine running a \texttt{select} loop over five typed channels:

\begin{lstlisting}
type CRDT struct {
    state      map[string]SettingEntry
    localCh    chan SettingEntry
    remoteCh   chan SettingEntry
    queryCh    chan queryRequest
    snapshotCh chan snapshotRequest
    gcCh       chan int64
}
\end{lstlisting}

\textbf{No mutex} is required: only the actor goroutine accesses channels and
CRDT state ensuring operation atomicity. External callers (\texttt{NotifyLocal}, \texttt{NotifyRemote},
\texttt{Snapshot}, \texttt{queryRequest}, \texttt{PurgeTombstones}) communicate exclusively through
channels. Snapshot and query requests are buffered synchronous response channels thus blocking
untill the request is serviced.

Callbacks (\texttt{OnBroadcast}, \texttt{OnFileSync}) are invoked
synchronously within the actor loop. The \texttt{OnFileSync} callback
dispatches disk writes on a separate goroutine to prevent blocking the actor loop
on I/O.

% ─────────────────────────────────────────────────────────────────────────────
\section{Network Layer}
\label{sec:network}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{gRPC Service}

The inter-node protocol exposes two RPCs over gRPC:

\begin{itemize}
  \item \textbf{SubscribeStateUpdates} (server-streaming): the caller receives
    \texttt{ServerStateUpdate} messages through a persistent stream every time
    the server calls \texttt{Broadcast}. This ensures real-time propagation of
    local edits.
  \item \textbf{Sync} (unary): the caller sends its local snapshot, the
    server updates its entries and responds with local server side entries that
    are newer. Used for anti-entropy reconciliation.
\end{itemize}

The server maintains a map of subscriber channels protected by a mutex.
\texttt{Broadcast} under read-lock retrieves the subscribed peers channels, then
sends the update to each channel in a non locking manner. Slow subscribers
drop updates relying on the next anti-entropy pass to recover.

\subsection{Abstraction Boundary}

A design decision was taken such all protobuf conversion
(\texttt{ToProto}, \texttt{FromProto}, \texttt{SnapshotToProto}) is
confined in the \texttt{network} package. The \texttt{Client} methods accept
and return domain types (\texttt{types.Snapshot}, \texttt{types.SettingEntry}).
Conversion to and from wire format is performed internally.

This was not in the original design: an earlier iteration of \texttt{Client.Sync} and \texttt(Client.Broadcast)
accepted \texttt{[]*userpb.SettingEntry} and \texttt{[]*userpb.ServerStateUpdate},
thus requiring the controller to perform the conversion before each call.
Refactor moved conversion dependency into network package, making the transport
layer a true black box.

\subsection{Reconnection and Keepalives}

\texttt{Subscribe} method in \texttt{Client} wraps the streaming RPC in
a retry loop: on any error, including server restart or network
interruption, it waits three seconds and tries to re-establish the stream, first of all
performing a Sync call to replay any entries missed during the outage. gRPC
keepalive parameters are configured to detect dead connections.

% ─────────────────────────────────────────────────────────────────────────────
\section{Anti-Entropy Reconciliation}
\label{sec:antientropy}
% ─────────────────────────────────────────────────────────────────────────────

Broadcast over gRPC is best-effort: a slow, offline or reconnecting subscriber
may miss updates. A (state-based) CRDT that relies solely on broadcast for
propagation is not eventually consistent in presence of message losses,
nodes that miss a broadcast permanently diverge.

Production CRDT deployments (Riak, Dynamo \cite{dynamo2007}) address this
with a background anti-entropy mechanism: periodic full-state comparison
identifies and repairs divergence independently of the broadcast path (\href{https://docs.riak.com/riak/kv/latest/learn/concepts/active-anti-entropy/index.html}{Riak AAE}).

This system implements a lightweight variant: \texttt{syncAllPeers}
is called periodically by the controller, this sends a local snapshot to
each peer via the Sync RPC, updating peer entries and also local entries on returned values.
This is sufficient for eventual consistency: even if every broadcast is dropped, the
state will converge within one anti-entropy interval.

There is an unimplemented enhancment regarding exchanges of unupdated KV pairs: currently
can happen that 2 nodes at the same instant send a Sync request, this will on each node take a snapshot
of local state and send it, then upon reciving the snapshot both nodes update their local entries, both send the updates
that consider necessary to the other that already had applied them through the received snapshot.
This is solvable but out of context for this project purpose. (eg. Jitter, follower/leader Sync, Merkle tree, Delta state ecc.)

\begin{lstlisting}
func (ctrl *Controller) syncAllPeers(ctx context.Context) {
    snap := ctrl.crdt.Snapshot()
    for _, client := range ctrl.clients {
        go client.Sync(ctx, snap, func(e SettingEntry) {
            ctrl.crdt.NotifyRemote(e)
        })
    }
}
\end{lstlisting}

A 60-second interval was chosen as default: bounding divergence to one
minute in worst-case broadcast loss and generating negligible network traffic
for the cluster size targeted.

% ─────────────────────────────────────────────────────────────────────────────
\section{Crash-Safe Persistence}
\label{sec:persistence}
% ─────────────────────────────────────────────────────────────────────────────

After each merge, CRDT serialises its state to \texttt{crdt\_state.json} found in the CRDT working directory. The naive approach
using \texttt{os.WriteFile} is not crash-safe:
if the process is killed after \texttt{WriteFile} had truncated the file and before
the kernel had flushed all pages to disk, the node restarts with a
corrupted state file.

The correct pattern we used is the atomic rename:

\begin{lstlisting}
func atomicWriteFile(path string, data []byte) error {
    tmp, err := os.CreateTemp(filepath.Dir(path), ".tmp_*")
    // ... write data to tmp, call tmp.Sync(), tmp.Close() ...
    return os.Rename(tmp.Name(), path)
}
\end{lstlisting}

Three properties make this safe:
\begin{enumerate}
  \item The temporary file is in the same directory of the target,
    thus the rename is within a single filesystem so, atomic
    at VFS layer on POSIX systems.
  \item \texttt{tmp.Sync()} flushes file data to storage before
    the rename, preventing data to be in the page cache during power-loss.
  \item \texttt{os.Rename} is atomic on POSIX: any reader will see either the
    old file or the new file, never a partially-written intermediate state.
\end{enumerate}

If the process crashes before the rename, the temporary file is
left in the directory and cleaned up during the next \texttt{crdt.Init()} call,
the original file remains intact.

If the crash happens after state is persisted, \texttt{crdt.Reconcile} function is called
one time in \texttt{controller.Run()} thus fully restoring local file integrity on startup.
This is a limiting factor since on an already initialized node, if edits are made on the
setting file, while the node is offline, these will be discarded in favour of last known state.

% ─────────────────────────────────────────────────────────────────────────────
\section{Garbage Collection}
\label{sec:gc}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Problem}

\textbf{Deleted entries} persist in CRDT state file indefinitely, this also
happens for \textbf{temporary files} generated by \texttt{atomicRead} upon system
crash. In systems with frequent delete-recreate cycles over large KV stores,
tombstone accumulation brings two problems: memory leak over time,
slower \texttt{Sync} and \texttt{crdt.Run} since larger snapshots are required.

\subsection{Init cleanup}

With regard to orphaned tmp files from \texttt{atomicRead}, the best solution in this case
was also the easiest. Removal of all pattern matched tmp files on \texttt{crtd.Init} call.

\subsection{TTL-Based Approach}

The system also implements a TTL-based garbage collector. A ticker in the
controller calls \texttt{crdt.PurgeTombstones(ttl)} passing the TTL.
The CRDT actor then processes the request from \texttt{gcCh} channel monitored
in its Run loop, thus maintaining the single-writer on \texttt{c.state}, removing
tombstones with wallTime older than \texttt{now - ttl}.

\begin{lstlisting}
func (c *CRDT) purgeTombstones(ttl int64) {
    diff := time.Now().UnixNano() - ttl
    purged := 0
    for key, entry := range c.state {
        if entry.Deleted && entry.Clock.WallTime < diff {
            delete(c.state, key)
            purged++
        }
    }
    if purged > 0 {
        c.saveState()
        log.Printf("tombstone GC: purged %d entries", purged)
    }
}
\end{lstlisting}

The default TTL is 14 days, twice the intended maximum node downtime
after which an operator would investigate. This ensures that tombstones had been
observed and applied by all healthy nodes via anti-entropy before it is removed.
After a week of downtime a node supposively needs a re-initiation.

\subsection{Safety Considerations}

The TTL approach has one known failure point: if a node is offline for longer
than the TTL and holds a non-deleted entry for a key tombstoned on all other nodes,
the node will republish entry thus the deletion will appear to be reversed.
This is solved by forcing the node to re-initialize with empty settings wether
on startup the node follows \texttt{time.Now() - lastShutdownTime > TTL}

There are also 2 known limitations we call sputtering cycle, determined by
clock skew on machines running a tombstone garbage collection, then
performing a Sync request. This will, from the skewed node, propagate again the GC'd entry.
Moreover if a Sync request is performed just before a GC message is processed
in the CRDT actor, Sync cycle brings with it GC'd entries creating a the same sputtering problem.
These are mitigated through wire-layer filtering in the \texttt{Sync} handler.
Before including a local entry in the response or applying an incoming remote entry,
the handler skips any entry where \texttt{Deleted = true} and
\texttt{now - wallTime > TTL}. Filtering is applied in both directions preventing resurrection
of already-collected tombstones regardless of clock skew or snapshot-before-GC ordering
between nodes.

% ─────────────────────────────────────────────────────────────────────────────
\section{Lifecycle and Error Handling}
\label{sec:lifecycle}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Init / New Separation}

The CRDT package exposes three functions:
\begin{itemize}
  \item \texttt{New(workDir)} instantiates a \texttt{*CRDT} value with
    buffered channels. Pure allocation, no I/O, no error.
  \item \texttt{Init()} loads an existing node: cleans up any orphaned
    atomic-write tmp files, reads \texttt{node\_id}, and deserialises
    \texttt{crdt\_state.json}. Returns an error if the working directory
    has not been initialised (\emph{``run init first''}).
  \item \texttt{InitNew(entries)} bootstraps a fresh node: creates the
    working directory, writes \texttt{node\_id} if absent, stamps all
    supplied entries with a fresh HLC tick, and persists the resulting
    state. Used both at first-time setup and when forcing a clean slate.
\end{itemize}

The controller decides which path to take via \texttt{fixNodeState}:

\begin{lstlisting}
func (ctrl *Controller) fixNodeState(lastShutdown int64) error {
    if lastShutdown != -1 &&
        time.Now().UnixNano()-lastShutdown > ctrl.tombstoneTTL {
        return ctrl.crdt.InitNew(types.Settings{})
    }
    return ctrl.crdt.Init()
}
\end{lstlisting}

\texttt{checkLastShutdown()} returns \texttt{-1} when no shutdown file
exists (first startup after \texttt{InitNew} or \texttt{InitEmptyNode}),
causing \texttt{Init()} to load the freshly written state. If the node was
offline longer than the tombstone TTL, \texttt{InitNew} reinitialises with
an empty state making the node re-sync from peers rather than republishing
entries already garbage collected cluster wide.
\texttt{saveShutdownTime()} is deferred in \texttt{Run} and records the
clean-shutdown timestamp used by the next invocation.

\subsection{Error Propagation Principle}

Functions in library packages (\texttt{crdt}, \texttt{network}) return
errors rather than calling \texttt{log.Fatal} or \texttt{os.Exit} directly.
\texttt{log.Fatal} belongs only at the application boundary (\texttt{main})
where there is sufficient context to decide that an error is truly
unrecoverable. This makes every package independently testable without the test process exiting.

% ─────────────────────────────────────────────────────────────────────────────
\section{Testing Strategy}
\label{sec:testing}
% ─────────────────────────────────────────────────────────────────────────────

\subsection{Clock Injection}

The \texttt{HLC} struct accepts a \texttt{Clock} interface via parametrized
options at construction time. Tests inject \texttt{MockClock} to control
wallTime advancement deterministically, covering edge cases such as
backward clock steps, simultaneous ticks, and the case where system clock
overtakes both local and received wallTime, resetting logical counter to zero.

\subsection{Network Tests with bufconn}

The \texttt{network} package is tested with \texttt{google.golang.org/grpc/test/bufconn},
an in-memory listener that presents a real gRPC connection without opening
any network socket. This allows tests to exercise the full gRPC stack without port
allocation or network related problems.

\subsection{Coverage}

Line coverage targets were established and maintained:
\texttt{types} at 100\%, \texttt{crdt} at approximately 90\%, \texttt{logic}
at approximately 90\%, \texttt{network} at approximately 91\%, and
\texttt{controller} at approximately 92\%. The remaining uncovered branches are structurally
untestable without process-level mocking, \texttt{json.Marshal} on \texttt{map[string]string} (the error path never
fires in the Go standard library), and the \texttt{fsnotify} closed-channel
path (fires only after the watcher goroutine has already exited).

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}
% ─────────────────────────────────────────────────────────────────────────────

This paper presented a production-quality CRDT-based configuration
synchronisation system. The core LWW CRDT provides convergence
under concurrent writes and network partitions. The actor concurrency model
eliminates lock contention. Anti-entropy reconciliation bounds
divergence to Sync interval regardless of broadcast reliability.
Atomic persistence through \texttt{os.Rename} prevents corruption on ungracefull shutdown.
Tombstone GC fixes CRDT state growth for long-running deployments.

Several design decisions evolved during implementation. The protobuf
abstraction boundary was moved into the \texttt{network} package, removing
transport-layer knowledge from the controller. Error propagation got modified
to return \texttt{error} from library code and keep
\texttt{log.Fatal} in the application entry-point. Forcefull init had ben added
for nodes rejoining the cluster after more than TTL elapsed since last shutdown. A full
reconciliation cycle had ben added upon start to make sure ungracefull shutdowns did not
produce stale entries.

The system does not address gossip-based broadcast efficiency, or metrics
observability, all known gaps appropriate for future work in production
deployments at larger scale.

% ─────────────────────────────────────────────────────────────────────────────
\begin{thebibliography}{00}

  \bibitem{shapiro2011}
  M.~Shapiro, N.~Pregui\c{c}a, C.~Baquero, and M.~Zawirski,
  ``A comprehensive study of convergent and commutative replicated data types,''
  INRIA, Tech. Rep. RR-7506, 2011.

  \bibitem{kulkarni2014}
  S.~S.~Kulkarni and M.~Demirbas,
  ``Logical physical clocks and consistent snapshots in globally distributed
  databases,''
  in \emph{Proc. 18th Int. Conf. Principles of Distributed Systems (OPODIS)},
  Cortina d'Ampezzo, Italy, Dec.\ 2014.

  \bibitem{dynamo2007}
  G.~DeCandia, D.~Hastorun, M.~Jampani, G.~Kakulapati, A.~Lakshman,
  A.~Pilchin, S.~Sivasubramanian, P.~Vosshall, and W.~Vogels,
  ``Dynamo: Amazon's highly available key-value store,''
  in \emph{Proc. 21st ACM Symp. Operating Systems Principles (SOSP)},
  Stevenson, WA, Oct.\ 2007, pp.\ 205--220.

\end{thebibliography}

\end{document}
